---
title: "Statistical Inference Course Project Part 2"
author: "Scott Lin"
date: "January 21, 2015"
output: pdf_document
---

## Description

This project is to analyze the ToothGrowth data set and run some hypothesis testing on the supplement and dosage. 

---

##Q1: Load the ToothGrowth Data Set

```{r}
library(datasets)
data(ToothGrowth) 
```

---

##Q2: Provide a basic summary of the data

From the following R Code:

```{r}
table(ToothGrowth$supp,ToothGrowth$dose)
```

I have figured out that the data frame includes 3 columns with 3 doses for 60 observations. That's 10 observation per combined supp/dose pair. It's not a paired test as there are no columns that describe the subject. I've combined the doses with the supp and made a box/whisker plot from it. 

```{r, echo=FALSE, fig.height=4, fig.width=6}
CToothGrowth <- data.frame(combined = paste(ToothGrowth$supp, ToothGrowth$dose), ToothGrowth)
CToothGrowth <- CToothGrowth[order(CToothGrowth[4], CToothGrowth[3]),]
plot(CToothGrowth$combined, CToothGrowth$len)
```

First level observations is that both drugs get more "potent" as the doses increase. VC seems to be weaker in lower doses but has a wider range for a dose of 2.0. My geusses would be that OJ outperforms VC at lower doses, but higher doses will probably be inconclusive as the box/whisker plots overlap completely.

---

##Q3: Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering). 

##State your assumptions:

I'm assuming that:

- I'm going to use a T-test due to small sample size of 10
- This is an unpaired test as there are no values indicating whot he subject of the tests are
- Given the small sample size, I'm assuming it's a subset of the data and we don't know the true variance of the population, so i'm going to assume all variances are unequal.
- Longer tooth is better
- The object, is to find the drug with the dosage that grows the longest teeth. 
- It's clear from the box/whisker plot, the teeth grows longer, the higher the dosage so I'll concentrate at comparing dosage for the drugs at every level

There are 3 dosage options so there will be 3 tests and I'm going to test each and one of them individually.

##Did VC 0.5 grow longer teeth than OJ 0.5? (H0 is VC 0.5):
```{r}
VC05 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5, "len"]
OJ05 <- ToothGrowth[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 0.5, "len"]
t.test(VC05,OJ05,paired=FALSE, var.equal=FALSE)
```

The p-value is at a significance level (0.006). The t value is -3.16 and the confidence level is below 0 completely. We can reject the null hypothesis and say OJ at dose 0.5 is better. 

##Did VC 1.0 grow longer teeth than OJ 1.0? (H0 is VC 1.0):
```{r}
VC10 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0, "len"]
OJ10 <- ToothGrowth[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 1.0, "len"]
t.test(VC10,OJ10,paired=FALSE, var.equal=FALSE)
```

The p-value is at a significance level (0.001). The t value is -4.03 and the confidence level is below 0 completely. We can reject the null hypothesis and say OJ at dose 1.0 is better. 

##Did VC 2.0 grow longer teeth than OJ 2.0? (H0 is VC 2.0):
```{r}
VC20 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2.0, "len"]
OJ20 <- ToothGrowth[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 2.0, "len"]
t.test(VC20,OJ20,paired=FALSE, var.equal=FALSE)
```

The p-value is at a insignificant level (0.96). The t value is 0.04 and the confidence level is surrounds 0 at -3.6 to 3.8. It's inconclusive at the 2.0 dosage level, and we can not reject the null hypothesis. 

---

##Q4: State your conclusions. (Assumptions stated above)

I conclude that for both VC and OJ, higher doses leads to longer teeth. For both 0.5 and 1.0, OJ is clearly better at delivering the vitamin C required for longer teeth. At the 2.0 level, it's inconclusive which one is better, but I would choose OJ due to its similar mean, and smaller standard deviation. 

---

##Appendix

Let's see if it's true that increasing dosage is better:

```{r}
t.test(VC05, VC10, paired=FALSE, var.equal=FALSE)
```

VC 1.0 is better than VC 0.5. 

```{r}
t.test(VC10, VC20, paired=FALSE, var.equal=FALSE)
```

VC 2.0 is better than VC 1.0

```{r}
t.test(OJ05, OJ10, paired=FALSE, var.equal=FALSE)
```

OJ 1.0 is better than OJ 0.5

```{r}
t.test(OJ10, OJ20, paired=FALSE, var.equal=FALSE)
```

OJ 2.0 is better than OJ 1.0, not by much though. The p-value is 0.04.